### 2019-01-25 personal-statement.tex
Updated personal statement

\documentclass[11pt,final,twoside]{article} %setup
%set margins and load symbols and such
\usepackage[top=20mm,inner=20mm,outer=30mm,bottom=20mm,
marginparwidth=23mm,marginparsep=4mm]{geometry}
%header of document
\title{Personal Statement}
\author{Grant Haataja}
\date{\today}
%set paragraph spacing
\linespread{1.5}

%make stuff happen
\begin{document}
	\maketitle

One of the most important questions one can ask themselves in life is, 'what do I find fulfilling?' For me, the answer is problem solving. Solving complicated problems in an elegant and efficient manner is one of the most satisfying things of life I have found.

That is why I have chosen to study Mathematics and Computer Science. Both are disciplines whose very foundation is built on problem solving. And my greatest interest is to study and research coding theory, which lies directly at the intersection of Math and Computer Science. With a strong background in both subjects, I feel that I can bring a broader perspective to my research.

I have paid my own way in life since I started attending Williston State College at age 16. My parents make enough money to prevent me from receiving financial aid, but not enough for them to help me pay for my school. I’ve worked summer internships for up to 100 hours a week, I’ve had jobs while taking up to 20 credits in a semester, and found every opportunity I could to make my dreams a reality.

I am currently working 2 jobs for a total of around 40 hours per week, while taking 16 credits and acting as co-lead for the software development team for Advanced Rocketry Club on campus. The US MASTER scholarship would help immensely to take some of my workload stress off my shoulders and allow me to focus more deeply on my classes and exciting research opportunities.

My career goals are to graduate with a double major and research experience, work in my field to learn real-world applications and provide service to society, and eventually return to graduate school and become a professor at a university so I can continue to pursue my passion for Math and research, as well as help other students reach their full potential.

\end{document}