### 2019-02-04 resume.tex
Built resume in LaTeX

\documentclass{article}

%call packages
\usepackage{titlesec}
\usepackage{titling}
\usepackage[margin=1.2in]{geometry}
\usepackage{hyperref}

%format titles for sections, no numbers and no indentation
\titleformat{\section}
{\huge\bfseries}
{}
{0em}
{}[\titlerule]

%format titles for subsections, bullet points and 0.3 indentation
\titleformat{\subsection}[runin]
{\bfseries\large}
{$\bullet$}
{0.3em}
{}

%format titles for subsubsections, bullet points and 0.3 indentation
\titleformat{\subsubsection}
{\bfseries\large}
{$\bullet$}
{0.3em}
{}

%format maketitle command
\renewcommand{\maketitle}{
\begin{center}
{\huge\bfseries
\theauthor}

\vspace{0.5em}

grant.haataja@und.edu

701-580-6305

\url{https://www.linkedin.com/in/grant-haataja-40526a144/}

\url{https://und.joinhandshake.com/users/10782099}

\url{https://github.com/grantHaataja?tab=repositories}

\end{center}
}

\begin{document}
\title{R\'esum\'e}
\author{Grant Haataja}
\maketitle

\section{Education}

\subsection{University of North Dakota:}
Junior year double major in Computer Science and Mathematics with cumulative GPA of 3.8

\subsection{Williston State College:}
Associate's Degree in Science with High Honors

\subsection{High School:}
Home-schooled, graduated at age 16 with 3.9 GPA

\section{Leadership Experience}

\subsection{Leadership:}
Software Development team co-lead for Advanced Rocketry Club

\subsection{Volunteering:}
Alerus Center - Fight the Frost

\section{Work Experience}

\subsubsection{Helix Wine \& Bites (01/18/19 - Current)} 
\textbf{Bartender/Server:} Providing exceptional food and drink service to customers

\subsubsection{UND Math Department (11/26/19 - Current)}
\textbf{Math Tutor:} Helping students learn math concepts such as college algebra, calculus, discrete math, differential equations, set theory and logic, and abstract algebra

\subsubsection{Advanced Engineering and Environmental Services (05/11/17 - 11/18/19)}
\textbf{Engineering Assistant:} Writing daily reports, taking construction photos, GPS data collection and organization, coordinating with contractor and client for smooth project development

\section{Skills}

\subsection{Technical Skills:}
Java, C, Python, Bash, {\LaTeX}, GitHub, Linux CLI

\subsection{General Skills:}
Customer service, communication, teamwork, organization skills

\end{document}