### 2019-02-05 resume.tex
Updated resume

\documentclass{article}

%call packages
\usepackage{titlesec}
\usepackage{titling}
\usepackage[margin=.9in, top=0.59in]{geometry}
\usepackage{hyperref}

%format titles for sections, no numbers and no indentation
\titleformat{\section}
{\huge\bfseries}
{}
{0em}
{}[\titlerule]

%format titles for subsections, bullet points and 0.3 indentation
\titleformat{\subsection}[runin]
{\bfseries\large}
{$\bullet$}
{0.3em}
{}

%format titles for subsubsections, bullet points and 0.3 indentation
\titleformat{\subsubsection}
{\bfseries\large}
{$\bullet$}
{0.3em}
{}

%format maketitle command
\renewcommand{\maketitle}{
\begin{center}
{\huge\bfseries
\theauthor}

\vspace{0.5em}

\large grant.haataja@und.edu

\large 701-580-6305

\end{center}
}

\begin{document}
\title{R\'esum\'e}
\author{Grant Haataja}
\maketitle

\section{Education}

\subsection{University of North Dakota:}
Double major in Computer Science and Mathematics with a current cumulative GPA of 3.80. Expected Graduation Date: 12/2020

\subsection{Williston State College:}
Associate's Degree in Science with High Honors. Graduated in May, 2016 with a cumulative GPA of 3.85

\section{Projects}

\subsection{Advanced Rocketry Club:}
Developing software on embedded computer chips to control the systems of a liquid-propelled rocket for the Base 11 Space Challenge

\subsection{Association of Computing Machinery:}
Participate in regional programming competitions

\section{Work Experience}

\subsubsection{UND Math Department (11/26/19 - Current)}
\textbf{Math Tutor:} Helping students learn math concepts such as college algebra, calculus, discrete math, differential equations, set theory and logic, and abstract algebra

\subsubsection{Helix Wine \& Bites (01/18/19 - Current)} 
\textbf{Bartender/Server:} Providing exceptional food and drink service to customers

\subsubsection{Advanced Engineering and Environmental Services (05/11/17 - 11/18/18)}
\textbf{Engineering Assistant:} 
Oversaw and managed 3 construction crews for a 16-mile transmission waterline. Provided quality control on behalf of the client, and coordinated smooth project development

\section{Skills}

\subsection{Technical Skills:}
Java, C, Python, Bash, GitHub, Linux shell, {\LaTeX}

\subsection{General Skills:}
Customer service, communication, teamwork, organization skills

\section{Links}

\subsection{}
\url{https://www.linkedin.com/in/grant-haataja-40526a144/}

\subsection{}
\url{https://und.joinhandshake.com/users/10782099?buildFromResumeId=9939568}

\subsection{}
\url{https://github.com/grantHaataja?tab=repositories}

\end{document}