### 2019-05-04 csci487/project/paper.tex
Worked on 487 paper

\documentclass[conference]{IEEEtran}
%\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{float}
\graphicspath{ {./} }
\usepackage{listings}
\usepackage{color}

\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}

\lstdefinestyle{mystyle}{
	backgroundcolor=\color{backcolour},   
	commentstyle=\color{codegreen},
	keywordstyle=\color{magenta},
	numberstyle=\tiny\color{codegray},
	stringstyle=\color{codepurple},
	basicstyle=\footnotesize,
	breakatwhitespace=false,         
	breaklines=true,                 
	captionpos=b,                    
	keepspaces=true,                 
	numbers=left,                    
	numbersep=5pt,                  
	showspaces=false,                
	showstringspaces=false,
	showtabs=false,                  
	tabsize=2
}

\lstset{style=mystyle}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{CSci487 Penetration Testing Project: AILEE}

\author{\IEEEauthorblockN{Grant Haataja}
\IEEEauthorblockA{\textit{UND Computer Science} \\
Grand Forks, USA \\
grant.haataja@und.edu}
\and
\IEEEauthorblockN{David Wilson}
\IEEEauthorblockA{\textit{UND Computer Science} \\
Grand Forks, USA \\
david.andrew.wilson@und.edu}
\and
\IEEEauthorblockN{Michael Turnbull}
\IEEEauthorblockA{\textit{UND Computer Science} \\
Grand Forks, USA \\
michael.turnbull@und.edu}
}

\maketitle

\begin{abstract}
This document details the planning, development, and workings of the penetration testing game AILEE, created as a final project for CSCI 487 Penetration Testing class at the University of North Dakota.
\end{abstract}

%\begin{IEEEkeywords}
%component, formatting, style, styling, insert
%\end{IEEEkeywords}

\section{Introduction}
For this project on penetration testing topics, a hacking simulation game was created. The premise of the game is as follows: the user plays the role of a penetration-testing AI software named AILEE, which stands for Artificial Intelligence Linux Exploit Environment. The game takes place exclusively in a Linux-style terminal environment, with a limited arsenal of commands for the player to use. As the player progresses through the game and “learns” as an AI, the commands available for use increase. Throughout the game, the player is given typed instructions and information from the AI’s administrator to assist in learning.

There are two targets to hack in this demo, although there is much potential for expansion. The game uses simulated port scanning, vulnerability scanning, exploitation, and other penetration testing tools to mimic real-life penetration testing methods. Additionally, the game features a storyline with three possible endings, depending on player actions. Special care was taken to handle proper sequence of events.

\section{Investigation}

\subsection{Planning the Project}

Before beginning the development of the game, a suitable platform to run the environment needed to be found. The website Repl.it was decided upon, due to their extensive language support and the ability for multiple people to work simultaneously and have all changes automatically saved to the cloud. [1] The “Multiplayer” mode, as this feature was called, still had a lot of bugs, so forking the project and saving work manually was still necessary, but overall it made the development of AILEE much smoother.

Python3 was selected as the programming language of choice, due to its ease of scripting and strong object-oriented nature. The various classes corresponding to different aspects of the game and environment would be programmed separately, as well as Python scripts for each command available to the player, and every storyline event that could be run. The original plan was for there to be three different targets for the player to hack, but due to time limitations the scope was decreased to two targets.

To enable smooth graphics for the intro screen and the game’s ending events, the Python Curses library was referenced and used extensively. [2] This provided the ability to control keyboard input while text displayed on the screen or the ending event graphics played, to increase the smoothness of gameplay.

\section{Project Description}

\subsection{Intro Screen}\label{AA}
For the graphics of the intro screen for the game, ASCII art was used to spell the word AILEE, along with a selection for New Game or Exit. The user can move between the selection using the up or down arrow keys and choose by pressing the enter key. Selecting Exit will cause the terminal session within Repl.it to exit and the game will have to be run again, selecting New Game creates a new session and runs the game.

In addition to these, pressing the up arrow six times in a row will show a hidden third selection, Skip Dialog. This will run the game without displaying any of the instructions and information from the administrator to AILEE, and was very useful for testing the game during development. This mode is not explained or mentioned in the game, as it is not recommended to play without reading the dialogue.

The intro screen makes use of the Python curses library to allow smooth use of the arrow keys keyboard input and prevent buggy graphics.

\subsection{Starting the Game}
Upon choosing New Game, the user watches as the administrator logs into their account and launches AILEE.exe to start a new shell. After the shell loads, the first event triggers and text displays on the screen to inform the user what is going on. The administrator gives a brief explanation, and then the user is free to experiment with the Linux-style terminal environment. The terminal runs in the Shell class, (in tandem with the Game and DoStory classes), which supports multiple terminals on various computers. The code for the Shell class is as follows:

\lstinputlisting[language=Python]{shell.py}

The user is encouraged to try out the various possible commands, which can be displayed using the \textit{help} command. The story continues after the user has ran ten commands, (they can be the same or different commands, it doesn't matter).

\subsection{Gameplay and Executables}
As the game progresses, the user will utilize the available penetration testing tools to gain access to the target computers. There are tools for finding IP addresses, port scanning, vulnerability scanning, exploitation, password cracking, and connecting through the ftp file sharing network. Generally, some of the commands require the results from running other commands in order to be run successfully.

In particular, the \textit{exploit} command is extremely useful for gaining access to other computers. This executable is similar to the Metasploit framework which is commonly used in real-life penetration testing. In the game, it only has two possible exploits to choose from, and if ran against a vulnerable IP address with the proper port selection, will successfully open a new shell on the target computer.

\lstinputlisting[language=Python]{exploit.py}

Above is the code for the \textit{exploit} executable in the game. In terms of options, it does not compare to the Metasploit framework, but the goal was to create it to feel similarly in the terminal environment.

\begin{figure}[htbp]
	\centerline{\includegraphics[scale=2]{exploit-pic}}
	\caption{Running the \textit{exploit} executable}
	\label{fig}
\end{figure}

First, the user enters the number corresponding to the exploit they wish to run. Then, the target IP address is entered, followed by the specific port, and finally the command 'run' must be entered and the exploitation software will attempt to gain access to the target (Fig. 1).

\begin{figure}[htbp]
	\centerline{\includegraphics[scale=2]{loading-new-shell}}
	\caption{Gaining a shell on target machine}
	\label{fig}
\end{figure}

If the exploit is successful, the screen will clear and the words "Loading new shell...." will appear on the screen with increasing dots as the shell loads (Fig 2).

\begin{figure}[htbp]
	\centerline{\includegraphics[scale=2]{access-granted}}
	\caption{A new shell on an exploited target}
	\label{fig}
\end{figure}

Once the shell loads, the name of the computer exploited will be shown before the /\$ symbol where commands are typed (Fig 3).

Most of the other executables play an important role in the game, with a few exceptions that were included for comedic value.

\subsection{\LaTeX-Specific Advice}

Please use ``soft'' (e.g., \verb|\eqref{Eq}|) cross references instead
of ``hard'' references (e.g., \verb|(1)|). That will make it possible
to combine sections, add equations, or change the order of figures or
citations without having to go through the file line by line.

Please don't use the \verb|{eqnarray}| equation environment. Use
\verb|{align}| or \verb|{IEEEeqnarray}| instead. The \verb|{eqnarray}|
environment leaves unsightly spaces around relation symbols.

Please note that the \verb|{subequations}| environment in {\LaTeX}
will increment the main equation counter even when there are no
equation numbers displayed. If you forget that, you might write an
article in which the equation numbers skip from (17) to (20), causing
the copy editors to wonder if you've discovered a new method of
counting.

{\BibTeX} does not work by magic. It doesn't get the bibliographic
data from thin air but from .bib files. If you use {\BibTeX} to produce a
bibliography you must send the .bib files.

{\LaTeX} can't read your mind. If you assign the same label to a
subsubsection and a table, you might find that Table I has been cross
referenced as Table IV-B3.

{\LaTeX} does not have precognitive abilities. If you put a
\verb|\label| command before the command that updates the counter it's
supposed to be using, the label will pick up the last counter to be
cross referenced instead. In particular, a \verb|\label| command
should not go before the caption of a figure or a table.

Do not use \verb|\nonumber| inside the \verb|{array}| environment. It
will not stop equation numbers inside \verb|{array}| (there won't be
any anyway) and it might stop a wanted equation number in the
surrounding equation.

\subsection{Some Common Mistakes}\label{SCM}
\begin{itemize}
\item The word ``data'' is plural, not singular.
\item The subscript for the permeability of vacuum $\mu_{0}$, and other common scientific constants, is zero with subscript formatting, not a lowercase letter ``o''.
\item In American English, commas, semicolons, periods, question and exclamation marks are located within quotation marks only when a complete thought or name is cited, such as a title or full quotation. When quotation marks are used, instead of a bold or italic typeface, to highlight a word or phrase, punctuation should appear outside of the quotation marks. A parenthetical phrase or statement at the end of a sentence is punctuated outside of the closing parenthesis (like this). (A parenthetical sentence is punctuated within the parentheses.)
\item A graph within a graph is an ``inset'', not an ``insert''. The word alternatively is preferred to the word ``alternately'' (unless you really mean something that alternates).
\item Do not use the word ``essentially'' to mean ``approximately'' or ``effectively''.
\item In your paper title, if the words ``that uses'' can accurately replace the word ``using'', capitalize the ``u''; if not, keep using lower-cased.
\item Be aware of the different meanings of the homophones ``affect'' and ``effect'', ``complement'' and ``compliment'', ``discreet'' and ``discrete'', ``principal'' and ``principle''.
\item Do not confuse ``imply'' and ``infer''.
\item The prefix ``non'' is not a word; it should be joined to the word it modifies, usually without a hyphen.
\item There is no period after the ``et'' in the Latin abbreviation ``et al.''.
\item The abbreviation ``i.e.'' means ``that is'', and the abbreviation ``e.g.'' means ``for example''.
\end{itemize}
An excellent style manual for science writers is \cite{b7}.

\subsection{Authors and Affiliations}
\textbf{The class file is designed for, but not limited to, six authors.} A 
minimum of one author is required for all conference articles. Author names 
should be listed starting from left to right and then moving down to the 
next line. This is the author sequence that will be used in future citations 
and by indexing services. Names should not be listed in columns nor group by 
affiliation. Please keep your affiliations as succinct as possible (for 
example, do not differentiate among departments of the same organization).

\subsection{Identify the Headings}
Headings, or heads, are organizational devices that guide the reader through 
your paper. There are two types: component heads and text heads.

Component heads identify the different components of your paper and are not 
topically subordinate to each other. Examples include Acknowledgments and 
References and, for these, the correct style to use is ``Heading 5''. Use 
``figure caption'' for your Figure captions, and ``table head'' for your 
table title. Run-in heads, such as ``Abstract'', will require you to apply a 
style (in this case, italic) in addition to the style provided by the drop 
down menu to differentiate the head from the text.

Text heads organize the topics on a relational, hierarchical basis. For 
example, the paper title is the primary text head because all subsequent 
material relates and elaborates on this one topic. If there are two or more 
sub-topics, the next level head (uppercase Roman numerals) should be used 
and, conversely, if there are not at least two sub-topics, then no subheads 
should be introduced.

\subsection{Figures and Tables}
\paragraph{Positioning Figures and Tables} Place figures and tables at the top and 
bottom of columns. Avoid placing them in the middle of columns. Large 
figures and tables may span across both columns. Figure captions should be 
below the figures; table heads should appear above the tables. Insert 
figures and tables after they are cited in the text. Use the abbreviation 
``Fig.~\ref{fig}'', even at the beginning of a sentence.

\begin{table}[htbp]
\caption{Table Type Styles}
\begin{center}
\begin{tabular}{|c|c|c|c|}
\hline
\textbf{Table}&\multicolumn{3}{|c|}{\textbf{Table Column Head}} \\
\cline{2-4} 
\textbf{Head} & \textbf{\textit{Table column subhead}}& \textbf{\textit{Subhead}}& \textbf{\textit{Subhead}} \\
\hline
copy& More table copy$^{\mathrm{a}}$& &  \\
\hline
\multicolumn{4}{l}{$^{\mathrm{a}}$Sample of a Table footnote.}
\end{tabular}
\label{tab1}
\end{center}
\end{table}

Figure Labels: Use 8 point Times New Roman for Figure labels. Use words 
rather than symbols or abbreviations when writing Figure axis labels to 
avoid confusing the reader. As an example, write the quantity 
``Magnetization'', or ``Magnetization, M'', not just ``M''. If including 
units in the label, present them within parentheses. Do not label axes only 
with units. In the example, write ``Magnetization (A/m)'' or ``Magnetization 
\{A[m(1)]\}'', not just ``A/m''. Do not label axes with a ratio of 
quantities and units. For example, write ``Temperature (K)'', not 
``Temperature/K''.

\section*{Acknowledgment}

The preferred spelling of the word ``acknowledgment'' in America is without 
an ``e'' after the ``g''. Avoid the stilted expression ``one of us (R. B. 
G.) thanks $\ldots$''. Instead, try ``R. B. G. thanks$\ldots$''. Put sponsor 
acknowledgments in the unnumbered footnote on the first page.

\section*{References}

Please number citations consecutively within brackets \cite{b1}. The 
sentence punctuation follows the bracket \cite{b2}. Refer simply to the reference 
number, as in \cite{b3}---do not use ``Ref. \cite{b3}'' or ``reference \cite{b3}'' except at 
the beginning of a sentence: ``Reference \cite{b3} was the first $\ldots$''

Number footnotes separately in superscripts. Place the actual footnote at 
the bottom of the column in which it was cited. Do not put footnotes in the 
abstract or reference list. Use letters for table footnotes.

Unless there are six authors or more give all authors' names; do not use 
``et al.''. Papers that have not been published, even if they have been 
submitted for publication, should be cited as ``unpublished'' \cite{b4}. Papers 
that have been accepted for publication should be cited as ``in press'' \cite{b5}. 
Capitalize only the first word in a paper title, except for proper nouns and 
element symbols.

For papers published in translation journals, please give the English 
citation first, followed by the original foreign-language citation \cite{b6}.

\begin{thebibliography}{00}
\bibitem{b1} Repl.it, “The world's leading online coding platform,” repl.it. [Online]. Available: https://repl.it/site/features. [Accessed: 02-May-2019].

\bibitem{b2} A. M. Kuchling and E. S. Raymond, “Curses Programming with Python¶,” Curses Programming with Python - Python 3.7.3 documentation. [Online]. Available: https://docs.python.org/3/howto/curses.html. [Accessed: 02-May-2019].

\end{thebibliography}
\vspace{12pt}
\color{red}
IEEE conference templates contain guidance text for composing and formatting conference papers. Please ensure that all template text is removed from your conference paper prior to submission to the conference. Failure to remove the template text from your paper may result in your paper not being published.

\end{document}